### 2014-06-08 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago_Fire: add two figures, update writeup

\section{Geometry}
\label{geom}
The structure was a 2 1/2 story residential building with balloon frame construction and a wood gable roofing system. The structure had a 17.15 m (56 ft) by 6.15 m (20 ft) footprint (cf. Fig.~(\ref{fig:simp_geom})). This results in a floor area of approximately 105 m$^2$ (1100 ft$^2$) per floor and a main living area of approximately 110 m$^2$ (2200 ft$^2$). Interior walls were constructed from wood frame dimensioned lumber with 1.27 cm (0.5 in) gypsum board over top. Additionally there was a half-story attic that was used to house the structure's hot water heater. There were three sets of stairs in the structure: one front stairwell that connected the first and second levels, one rear stairwell that connected the first and second levels and a second rear set of stairs which provided access to the attic. Entry to the structure can be made either in the front (Alpha side) or rear (Charlie side). The fire, which is described in Section~\ref{fire}, started in the attic and spread to the rear porch (cf. Fig.~(\ref{fig:simp_geom})). Discussion of the fire's impact on interior conditions can be found in Section~\ref{results}. Fig.~(\ref{fig:alpha_ex}) shows the post-incident exterior of the alpha side of the structure. Note that the majority of the damage is to the structure's attic.

@@

@@

\section{Pressure}
As the fire grows in the attic and back porch, the pressure rises. Fig.~(\ref{fig:pres_159s}) shows the pressure rise in the structure just before the interior door failed in the simulation. On the rear porch, the pressure rise ranges from a 4 Pa over pressure 1 m off the floor to greater than 10 Pa at the ceiling.
\begin{figure}[!ht]
\includegraphics[width=.75\textwidth]{../Figures/west_50th_baseline_pres}
\input{../Scripts/colorbar_metric_pres}
\caption{FDS predicted pressure contours, 1 second prior to interior door failure.}
\label{fig:pres_159s}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_175}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature from baseline simulation, 15 seconds after interior door fails.}
\label{fig:temp_175s}
\end{figure}

>>>>

\section{Geometry}
\label{geom}
The structure was a 2 1/2 story residential building with balloon frame construction and a wood gable roofing system. The structure had a 17.15 m (56 ft) by 6.15 m (20 ft) footprint (cf. Fig.~(\ref{fig:simp_geom})). This results in a floor area of approximately 105 m$^2$ (1100 ft$^2$) per floor and a main living area of approximately 110 m$^2$ (2200 ft$^2$). Interior walls were constructed from wood frame dimensioned lumber with 1.27 cm (0.5 in) gypsum board over top. Additionally there was a half-story attic that was used to house the structure's hot water heater. There were three sets of stairs in the structure: one front stairwell that connected the first and second levels, one rear stairwell that connected the first and second levels and a second rear set of stairs which provided access to the attic. Entry to the structure can be made either in the front (Alpha side) or rear (Charlie side), see Fig.~(\ref{fig:simp_geom}). The fire, which is described in Section~\ref{fire}, started in the attic and spread to the rear porch. Discussion of the fire's impact on interior conditions can be found in Section~\ref{results}. Fig.~(\ref{fig:alpha_ex}) shows the post-incident exterior of the alpha side of the structure. Note that the majority of the damage is to the structure's attic.

@@

\section{Timeline}
\label{timeline}
From the NIOSH report~\cite{NIOSH:Bowyer}, a timeline of events was able to be created, see Table~\ref{tab:fire_info}. Due to a lack of information about when the fire ignited, an ignition time/start of the simulation needed to be estimated.  We estimate that the simulation starts at 17:19:20. Since the actually time was unknown, the fire growth rate in the computational model is such that when firefighters enter the structure, the hazard level in the model is similar to the real situation. A second unknown that was estimated was the time when the second floor door fails (2:40 into the simulation). While this time is not exact, the takeaway from the analysis in the following sections should be impact of the door's failure on the interior conditions of the structure. Table~\ref{tab:firesim} shows the fireground timeline along with the corresponding simulation times.

\begin{table}
\centering
\captionof{table}{Fire - Simulation Event Timeline}\label{tab:firesim}
\begin{tabular}{l l l}
\toprule[1.5pt]
Incident Time & Simulation Time & Fire Behavior / Fireground Operation  \\
\midrule
\multirow{2}{*}{17:15} &    & \multirow{2}{*}{\parbox{10cm} {Reported working fire with heavy smoke coming from the front and rear of the structure's attic.}} \\ 
         & & \\[.25cm] %blank rows exist for wrapping text. # of blanks is 1 less than number of rows
17:16 &      & Dispatch for a Still Alarm. \\ [0.25cm]
         &  \multirow{2}{*} {0:00} &  \multirow{2}{*}{\parbox{10cm} {Simulation starts. Attic fire ignites. Initial self ventilation on charlie side. Front door open.}} \\
         & & \\[0.25cm]
\multirow{2}{*}{17:20} & \multirow{2}{*}{0:40}  & \multirow{2}{*}{\parbox{10cm} {Attic self ventilation - heavy smoke out of front attic window and heavy smoke and fire out of rear.}} \\
         & & \\[.25cm] 
         & 0:45 & Model ventilation on charlie side. \\[.25cm] 
         & 0:55 & Model ventilation on beta side. \\[.25cm] 
17:21:00 &  & E123 Capt. and FF on 2$^{nd}$ floor with 1 3/4-in hose line. \\[.25cm]
17:21:30 & 2:10 & 1$^{st}$ floor exterior porch door opened. Attic fire ignites in model. \\[.25cm]
        & 2:15 & Additional model ventilation on beta side. \\[.25cm] 
        & 2:40 & Top of interior door fails in model. \\[.25cm] 
\multirow{2}{*}{17:23}    & \multirow{2}{*}{3:40} & \multirow{2}{*}{\parbox{10cm} {Heavy fire report in rear stairwell and enclosed porch. No response from E123: Capt. or FF.}} \\
         & & \\[.25cm]
17:24    &   & Water applied to attic from the rear exterior position. \\[.25cm]
17:27    & 5:00  & IC calls a Mayday. Simulation ends. \\[.25cm]
17:29    &   & E123 Capt. carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\section{Pressure}
As the fire grows in the attic and back porch, the pressure rises. Fig.~(\ref{fig:pres_159s}) shows the pressure rise in the structure just before the interior door failed in the simulation. On the rear porch, the pressure rise ranges from a 4 Pa over pressure 1 m off the floor to greater than 10 Pa at the ceiling.
\begin{figure}[!ht]
\includegraphics[width=.75\textwidth]{../Figures/west_50th_baseline_pres2}
\input{../Scripts/colorbar_metric_pres}
\caption{FDS predicted pressure contours, 1 second prior to interior door failure.}
\label{fig:pres_159s}
\end{figure}
Gases at elevated pressure will always travel toward an area of lower pressure. When the interior door fails, the lower pressure outlet is through the front door on the alpha side of the structure. As a result, the gases produced by the fire will flow through the interior hallway, through the kitchen, down the front stairs and out the front door. The velocity vectors in Fig.~(\ref{fig:velo_175s}), 15 seconds after the interior door fails, confirms this flow direction within the structure.
\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_velo_175}
\input{../Scripts/colorbar_std_velo}
\input{../Scripts/colorbar_metric_velo}
\caption{FDS predicted velocity vectors, 15 second after interior door failure.}
\label{fig:velo_175s}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[width=.675\textwidth]{../Figures/west_50th_baseline_175}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{FDS predicted temperature contours, 15 seconds after interior door fails.}
\label{fig:temp_175s}
\end{figure}
